---
title: "Analysis of the World Happiness Report"
author: "Anna Wallace"
institute: "SIS 750: Data Analysis"
date: "April 7, 2023"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.width = 7, fig.asp = 0.618, fig.align = 'center', out.width = '70%'
)
library(tidyverse)
library(dplyr)
library(kableExtra)
library(patchwork)
library(scales)
library(knitr)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(base_color = "#43418A",
          # fonts
          header_font_google = google_font("Martel"),
          text_font_google = google_font("Lato"),
          code_font_google = google_font("Fira Mono"))
```

---
# The Dataset

World Happiness Report, 2005-Present
- **Author:** Usama Battar on Kaggle
- **Source:** [Sustainable Development Solutions Network](https://www.kaggle.com/datasets/usamabuttar/world-happiness-report-2005-present)
- **Methodology:** Gallup World Poll data

--

This dataset has **2,199** observations of **13** variables:
- Life Ladder
- Corruption
- Social Support

--

New variable addition: `region`
---

### Graph 1. Life Evaluation per Level of Social Support 
```{r, include=FALSE} 
df = read_csv('World Happiness Report.csv', na = 'NA')

# filter by year and rename columns 

df1 = df %>%
  filter(Year == 2021) %>%
  rename(country = "Country Name",
         regionID = "Regional Indicator",
         life_ladder = "Life Ladder",
         gdp_per_cap = "Log GDP Per Capita",
         soc_support = "Social Support",
         healthy_life = "Healthy Life Expectancy At Birth",
         freedom = "Freedom To Make Life Choices",
         generosity = Generosity,
         corruption = "Perceptions Of Corruption",
         govt_confidence = "Confidence In National Government") %>%
  select(-"Positive Affect", -"Negative Affect") %>%
  mutate(
    region = case_when(regionID == "Central and Eastern Europe" ~ 'Europe',
                          regionID == "Commonwealth of Independent States" ~ 'Asia',
                          regionID == "East Asia" ~ 'Asia',
                          regionID == "Latin America and Caribbean" ~ 'North America',
                          regionID == "Middle East and North Africa" ~ 'MENA',
                          regionID == "North America and ANZ" ~ 'North America',
                          regionID == "South Asia" ~ 'Asia',
                          regionID == "Southeast Asia" ~ 'Asia',
                          regionID == "Sub-Saharan Africa" ~ 'Africa',
                          regionID == "Western Europe" ~ 'Europe')) %>%
  relocate(region, .before = regionID)

df1 = df1[complete.cases(df1$region), ]
df1$region = as.factor(df1$region)

# VIZ1
plot1 = df1 %>%
  ggplot(aes(x = life_ladder, y = soc_support, shape = region, color = region))+
  geom_point(size = 3) +
  labs(x = 'Life Ladder Rank',
       y = 'Social Support Level') +
  theme_minimal() +
  theme(
    axis.title.y = element_text(face = 'bold', color = 'mediumpurple4'),
    axis.title.x = element_text(face = 'bold', color = 'mediumpurple4'),
    legend.position = 'top',
    legend.title = element_text(face = "bold", color = 'mediumpurple4'))
```

```{r plot1, out.height = '400px', out.width = '600px'}
plot1
```
---
### Graph 2. Freedom of Life Choices per Perceived Government Corruption 


```{r, include=FALSE}
# create second VIZ - freedom to make choices by govt corruption
p2 = df1 %>%
  ggplot(aes(x = freedom, y = corruption)) +
  geom_point(size = 2, color = 'mediumpurple4') +
  geom_smooth(se=F, color='red')+
  labs(x = 'Freedom in Life Choices',
       y = 'Corruption') +
  theme_light()+
  theme(
    axis.title.y = element_text(face = 'bold', color = 'mediumpurple4'),
    axis.title.x = element_text(face = 'bold', color = 'mediumpurple4'))
  
# facet wrap by region
p3 = df1 %>%
  ggplot(aes(x=freedom, y=corruption)) +
  geom_point(color = 'mediumpurple4', alpha = (1/2), size = 3) +
  facet_wrap(~ region) +
  geom_smooth(lwd = 1.2, se=F, color='red') +
  labs(x = 'Freedom in Life Choices',
       y = 'Corruption') +
  theme_light()+
  theme(
    axis.title.y = element_text(face = 'bold', color = 'mediumpurple4'),
    axis.title.x = element_text(face = 'bold', color = 'mediumpurple4'),
    strip.background = element_rect(fill = "mediumpurple4"))
```

```{r plot2, echo=FALSE, message=FALSE, warning=FALSE, out.height = '450px', out.width = '600px' }
p2 / p3
```

---
class: inverse, middle, center

# Questions?